---
title: "Poverty"
author: "Febriany Lete"
date: "2024-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(haven)
library(foreign)
library(dplyr)
library(tidyr)
library(labelled)
library(ggplot2)
library(ggrepel)
```

Load Datasets

```{r, echo=FALSE}
kp3_202303 <- read_sav("kp43.sav")
kor1_202303 <- read_sav("kor_ind1.sav")
kor2_202303 <- read.dbf("53_ssn_202303_kor_ind2.dbf")
korrt_202303 <- read.dbf("53_ssn_202303_kor_rt_rev.dbf")

kp3_202303 <- kp3_202303 %>%
  mutate(GK = case_when(
    r101 == 53 & r105 == 1 ~ 614436,
    r101 == 53 & r105 == 2 ~ 471502,
     TRUE ~ NA_real_ 
  ))

kp3_202303 <- kp3_202303 %>%
  mutate(poverty = ifelse(kapita < GK, 1, 0))

kp3_202303$poverty <- factor(kp3_202303$poverty, levels = c(0, 1), labels = c("Tidak miskin", "Miskin"))

merged_kor1 <- merge(kor1_202303, kp3_202303, 
                     by = c("D_R", "urut", "r101", "r102", "r105", "psu", "ssu", "wi1","wi2"), 
                     all = TRUE)

korrt_202303 <- korrt_202303 %>%
  rename(urut = URUT, r101 = R101, r102 = R102, r105 = R105, psu = PSU, ssu = SSU, wi1 = WI1, wi2 = WI2)
merged_kor2 <- merge(korrt_202303, kp3_202303, 
                     by = c("urut", "r101", "r102", "r105", "psu", "ssu", "wi1","wi2"), 
                     all = TRUE)

```

## Karakteristik Sosial Demografi Rumah Tangga Miskin dan Tidak Miskin Menurut Wilayah

**Poverty line and population living in poverty**

Garis kemiskinan di NTT berdasarkan Susenas Maret 2023 adalah

\- Perkotaan: Rp. 614,436

\- Perdesaan: Rp. 471,502

```{r, echo=FALSE}

#kp3_202303$poverty <- factor(kp3_202303$poverty, levels = c(0, 1), labels = c("Tidak miskin", "Miskin"))

```

### 1. Average household size


```{r, echo=FALSE}
expend_jart <- kp3_202303 %>%
  group_by(poverty, r105) %>%
  summarize(mean_r301 = weighted.mean(r301, w = wert, na.rm = TRUE), 
            .groups = 'drop')
expend_jart$r105 <- factor(expend_jart$r105, 
                           levels = c(1, 2), 
                           labels = c("Perkotaan", "Perdesaan"))

expend_jartall <- kp3_202303 %>%
  group_by(poverty) %>%
   summarize(mean_r301 = weighted.mean(r301, w = wert, na.rm = TRUE), 
            .groups = 'drop')

combined_jart <- bind_rows(expend_jart, expend_jartall)
combined_jart <- combined_jart %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))


jart <- ggplot(combined_jart, aes(x = r105, y = mean_r301, fill = poverty)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(mean_r301, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            color = "black") +
  labs(fill = "Status Kemiskinan",
       x = "",
    y = "Rata-rata Jumlah Anggota Rumah Tangga"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "salmon"))
print(jart)
```

### 2. Education

Head of household education


```{r, echo=FALSE}
merged_kor1 <- merged_kor1 %>%
  mutate(didik5kat = case_when(
    r407 >= 5 & r614 == 25 ~ 1,
    r407 >= 5 & r614 %in% 1:5 ~ 2,
    r407 >= 5 & r614 %in% 6:10 ~ 3,
    r407 >= 5 & r614 %in% 11:17 ~ 4,
    r407 >= 5 & r614 %in% 18:24 ~ 5,
    TRUE ~ NA_real_ 
  ))
merged_kor1 <- merged_kor1 %>%
  mutate(didik5kat = factor(didik5kat,
                             levels = 1:5,
                             labels = c("Tidak Tamat SD", "SD", "SMP", "SMA", "Perguruan Tinggi"))) %>%
  set_variable_labels(didik5kat = "Pendidikan Tertinggi")


krt_edu <- merged_kor1 %>%
  filter(r403 == 1) %>%
  filter(!is.na(didik5kat)) %>%
  count(r105, poverty, didik5kat, wt = wert) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()


print(krt_edu)

krt_eduall <- merged_kor1 %>%
  filter(r403 == 1) %>%
  filter(!is.na(didik5kat)) %>%
  count(poverty, didik5kat, wt = wert) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

combined_edu <- bind_rows(krt_edu, krt_eduall)
combined_edu$r105 <- factor(combined_edu$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_edu <- combined_edu %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))


edu_chart <- ggplot(combined_edu, aes(x = r105, y = percentage, fill = didik5kat)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.6, 
            size = 2,
            color = "black") +
  labs(x = "",
       y = "Persentase",
       fill = "Tingkat Pendidikan") +
  theme(plot.title = element_text(hjust = 0.5))

print(edu_chart)
```

### 3. Housing

#### a) Home ownership

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(R1802_label = case_when(
    R1802 == 1 ~ "Milik Sendiri",
    R1802 == 2 ~ "Kontrak/Sewa",
    R1802 >= 3 ~ "Lainnya",
    TRUE ~ "Unknown"
  )) %>%
  mutate(R1802_label = factor(R1802_label, levels = c("Milik Sendiri", "Kontrak/Sewa", "Lainnya")))

ownership_reg <- merged_kor2 %>%
  filter(!is.na(R1802)) %>%
  count(r105, poverty, R1802_label, wt = wert) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(ownership_reg)

ownership_all <- merged_kor2 %>%
  filter(!is.na(R1802)) %>%
  count(poverty, R1802_label, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(ownership_all)

combined_ownership <- bind_rows(ownership_reg, ownership_all)
combined_ownership$r105 <- factor(combined_ownership$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_ownership <- combined_ownership %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

own_chart <- ggplot(combined_ownership, aes(x = r105, y = percentage, fill = R1802_label)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(x = "",
       y = "Persentase",
       fill = "Status Kepemilikan") +
  theme(plot.title = element_text(hjust = 0.5))

print(own_chart)
```

#### b) Dwelling Floor Area

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(R1804_new = case_when(
    R1804 <= 19 ~ "<= 19",
    R1804 %in% 20:49 ~ "20-49",
    R1804 %in% 50:99 ~ "50-99",
    R1804 %in% 100:149 ~ "100-149",
    R1804 >= 150 ~ ">=150",
    TRUE ~ "Unknown"
  )) %>%
  mutate(R1804_new = factor(R1804_new, levels = c("<= 19", "20-49", "50-99", "100-149", ">=150")))


floor_reg <- merged_kor2 %>%
  filter(!is.na(R1804_new)) %>%
  count(r105, poverty, R1804_new, wt = wert) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(ownership_reg)

floor_all <- merged_kor2 %>%
  filter(!is.na(R1804_new)) %>%
  count(poverty, R1804_new, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(ownership_all)

combined_floor <- bind_rows(floor_reg, floor_all)
combined_floor$r105 <- factor(combined_floor$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_floor <- combined_floor %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

floor_chart <- ggplot(combined_floor, aes(x = r105, y = percentage, fill = R1804_new)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -1, 
            size = 3,
            color = "black") +
  labs(x = "",
       y = "Persentase",
       fill = "Luas Lantai") +
  theme(plot.title = element_text(hjust = 0.5))

print(floor_chart)

```

#### c) Dwelling Floor Area per Capita

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(area = R1804 / R301,
         floor_cap = case_when(
           area <= 8 ~ 1,
           area > 8 & area <= 15 ~ 2,
           area > 15 ~ 3
         ))

merged_kor2 <- merged_kor2 %>%
  mutate(floor_cap = factor(floor_cap, 
                             levels = c(1, 2, 3),
                             labels = c('<= 8', '8-15', '> 15')))

floorcap_reg <- merged_kor2 %>%
  filter(!is.na(floor_cap)) %>%
  count(r105, poverty, floor_cap, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(floorcap_reg)

floorcap_all <- merged_kor2 %>%
  filter(!is.na(floor_cap)) %>%
  count(poverty, floor_cap, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(floorcap_all)

combined_floorcap <- bind_rows(floorcap_reg, floorcap_all)
combined_floorcap$r105 <- factor(combined_floorcap$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_floorcap <- combined_floorcap %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

floorcap_chart <- ggplot(combined_floorcap, aes(x = r105, y = percentage, fill = floor_cap)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -1, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "Luas Lantai per Kapita") +
  theme(plot.title = element_text(hjust = 0.5))

print(floorcap_chart)
```

#### d) Main material of the floor

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(floor = case_when(
    R1808 == 8 ~ 2,
    TRUE ~ 1
  ))
merged_kor2 <- merged_kor2 %>%
  mutate(floor = factor(floor, 
         levels = c(1, 2),
         labels = c('Bukan Tanah', 'Tanah')))

floor_reg <- merged_kor2 %>%
  filter(!is.na(floor)) %>%
  count(r105, poverty, floor, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(floor_reg)

floor_all <- merged_kor2 %>%
  filter(!is.na(floor)) %>%
  count(poverty, floor, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(floor_all)

combined_floor <- bind_rows(floor_reg, floor_all)
combined_floor$r105 <- factor(combined_floor$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_floor <- combined_floor %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

floor_chart <- ggplot(combined_floor, aes(x = r105, y = percentage, fill = floor)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "Jenis lantai") +
  theme(plot.title = element_text(hjust = 0.5))

print(floor_chart)
```

#### e) The main material of the widest wall

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
 mutate(wall = case_when(
    R1807 == 1 ~ 1,
    R1807 %in% c(3, 5) ~ 2,
    R1807 %in% c(2, 4, 6) ~ 3,
    TRUE ~ 4
  ))
merged_kor2 <- merged_kor2 %>%
  mutate(wall = factor(wall, 
         levels = c(1, 2, 3, 4),
         labels = c('Tembok', 'Kayu', 'Bambu', 'Lainnya')))

wall_reg <- merged_kor2 %>%
  filter(!is.na(wall)) %>%
  count(r105, poverty, wall, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(wall_reg)

wall_all <- merged_kor2 %>%
  filter(!is.na(wall)) %>%
  count(poverty, wall, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(wall_all)

combined_wall <- bind_rows(wall_reg, wall_all)
combined_wall$r105 <- factor(combined_wall$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_wall <- combined_wall %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

wall_chart <- ggplot(combined_wall, aes(x = r105, y = percentage, fill = wall)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "Jenis Tembok") +
  theme(plot.title = element_text(hjust = 0.5))

print(wall_chart)
```

#### f) Main material of the widest roof

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
 mutate(roof = case_when(
    R1806 %in% c(1, 2, 6) ~ 1,
    R1806 %in% c(3, 4) ~ 2,
    R1806 == 7 ~ 3,        
    R1806 %in% c(5, 8) ~ 4 
  ))
merged_kor2 <- merged_kor2 %>%
  mutate(roof = factor(roof, 
         levels = c(1, 2, 3, 4),
         labels = c('Beton/ Genteng/ Sirap', 'Seng/ Asbes', 'Ijuk/ Rumbia', 'Lainnya')))

roof_reg <- merged_kor2 %>%
  filter(!is.na(roof)) %>%
  count(r105, poverty, roof, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(roof_reg)

roof_all <- merged_kor2 %>%
  filter(!is.na(roof)) %>%
  count(poverty, roof, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(roof_all)

combined_roof <- bind_rows(roof_reg, roof_all)
combined_roof$r105 <- factor(combined_roof$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_roof <- combined_roof %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

roof_chart <- ggplot(combined_roof, aes(x = r105, y = percentage, fill = roof)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "Jenis Atap") +
  theme(plot.title = element_text(hjust = 0.5))

print(roof_chart)
```

#### g) Restroom facilities used by the household

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(restroom = case_when(
    R1809A == 1 ~ 1,
    R1809A %in% 2:3 ~ 2,
    TRUE ~ 3 
  ))

merged_kor2 <- merged_kor2 %>%
  mutate(restroom = factor(restroom, 
                         levels = c(1, 2, 3),
                         labels = c('Jamban Sendiri', 'Jamban Bersama/Komunal', 'Jamban Umum/ Tidak ada')))

restroom_reg <- merged_kor2 %>%
  filter(!is.na(restroom)) %>%
  count(r105, poverty, restroom, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(restroom_reg)

restroom_all <- merged_kor2 %>%
  filter(!is.na(restroom)) %>%
  count(poverty, restroom, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(restroom_all)

combined_restroom <- bind_rows(restroom_reg, restroom_all)
combined_restroom$r105 <- factor(combined_restroom$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_restroom <- combined_restroom %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

restroom_chart <- ggplot(combined_restroom, aes(x = r105, y = percentage, fill = restroom)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "Fasilitas") +
  theme(plot.title = element_text(hjust = 0.5))

print(restroom_chart)

```

#### h) Safe drinking water consumed by household

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(safe_dw = case_when(
    R1810A %in% c(3, 4, 5, 7, 10) ~ 1,
    R1810A %in% c(1, 2) & R1814A %in% c(3, 4, 5, 7, 10) ~ 1,
    TRUE ~ 0
  ))

merged_kor2 <- merged_kor2 %>%
  mutate(safe_dw = factor(safe_dw, 
                            levels = c(0, 1),
                            labels = c("Tidak ada akses", "Ada akses")))

safe_dw_reg <- merged_kor2 %>%
  filter(!is.na(safe_dw)) %>%
  count(r105, poverty, safe_dw, wt = FWT) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(safe_dw_reg)

safe_dw_all <- merged_kor2 %>%
  filter(!is.na(safe_dw)) %>%
  count(poverty, safe_dw, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(safe_dw_all)

combined_safe_dw <- bind_rows(safe_dw_reg, safe_dw_all)
combined_safe_dw$r105 <- factor(combined_safe_dw$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_safe_dw <- combined_safe_dw %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

safe_dw_chart <- ggplot(combined_safe_dw, aes(x = r105, y = percentage, fill = safe_dw)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "") +
  theme(plot.title = element_text(hjust = 0.5))

print(safe_dw_chart)
```

#### i) Source of Lighting

```{r, echo=FALSE}
merged_kor2 <- merged_kor2 %>%
  mutate(R1816_label = case_when(
    R1816 <= 3 ~ "Listrik",
    R1816 == 4 ~ "Bukan Listrik",
    TRUE ~ "Unknown"
  )) %>%
  mutate(R1816_label = factor(R1816_label, levels = c("Listrik", "Bukan Listrik")))

lighting_reg <- merged_kor2 %>%
  filter(!is.na(R1816)) %>%
  count(r105, poverty, R1816_label, wt = wert) %>%
  group_by(r105, poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(lighting_reg)

lighting_all <- merged_kor2 %>%
  filter(!is.na(R1816)) %>%
  count(poverty, R1816_label, wt = FWT) %>%
  group_by(poverty) %>%
  mutate(total_count = sum(n)) %>%
  mutate(percentage = round((n / total_count) * 100, 2)) %>%
  rename(count = n) %>%
  ungroup()

print(lighting_all)

combined_lighting <- bind_rows(lighting_reg, lighting_all)
combined_lighting$r105 <- factor(combined_lighting$r105, levels = c(1, 2), labels = c("Perkotaan", "Perdesaan"))
combined_lighting <- combined_lighting %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

lighting_chart <- ggplot(combined_lighting, aes(x = r105, y = percentage, fill = R1816_label)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) + 
  facet_wrap(~poverty) +
  theme_minimal() +
  geom_text(aes(label = round(percentage, 2)), 
            position = position_dodge(width = 1), 
            vjust = -0.5, 
            size = 3,
            color = "black") +
  labs(
       x = "",
       y = "Persentase",
       fill = "") +
  theme(plot.title = element_text(hjust = 0.5))

print(lighting_chart)
```


### 4. Consumption

#### a) Average Daily Per Capita Protein Consumption
```{r, echo=FALSE}
protein <- kp3_202303 %>%
  group_by(poverty, r105) %>%
  summarize(mean_protein = weighted.mean(prote_kap, w = wert, na.rm = TRUE), 
            .groups = 'drop')
protein$r105 <- factor(protein$r105, 
                           levels = c(1, 2), 
                           labels = c("Perkotaan", "Perdesaan"))

protein_all <- kp3_202303 %>%
  group_by(poverty) %>%
  summarize(mean_protein = weighted.mean(prote_kap, w = wert, na.rm = TRUE), 
            .groups = 'drop')

combined_protein <- bind_rows(protein, protein_all)
combined_protein <- combined_protein %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

protein_chart <- ggplot(combined_protein, aes(x = r105, y = mean_protein, fill = poverty)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(mean_protein, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            color = "black") +
  labs(
    x = "",
    y = "Konsumsi Protein Perkapita",
    fill = "Status Kemiskinan"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "salmon"))
print(protein_chart)
```

#### b)The distribution

```{r, echo=FALSE}
protein_scatter <- ggplot(kp3_202303, aes(x = expend, 
                                          y = prote_kap, 
                                          color = poverty)) +
  geom_jitter(width = 0.2, alpha = 0.6) +
  labs(
    x = "Rata-rata Pengeluaran Rumah Tangga (Rupiah)",
    y = "Konsumsi Kalori Perkapita",
    fill = "Status Kemiskinan"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("skyblue", "salmon"))

print(protein_scatter)

```


#### c) Average Daily Per Capita Calorie Consumption

```{r,  echo=FALSE}
calorie <- kp3_202303 %>%
  group_by(poverty, r105) %>%
  summarize(mean_calorie = weighted.mean(kalori_kap, w = wert, na.rm = TRUE), 
            .groups = 'drop')
calorie$r105 <- factor(calorie$r105, 
                           levels = c(1, 2), 
                           labels = c("Perkotaan", "Perdesaan"))

calorie_all <- kp3_202303 %>%
  group_by(poverty) %>%
  summarize(mean_calorie = weighted.mean(kalori_kap, w = wert, na.rm = TRUE), 
            .groups = 'drop')

combined_calorie <- bind_rows(calorie, calorie_all)
combined_calorie <- combined_calorie %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

cal_chart <- ggplot(combined_calorie, aes(x = r105, y = mean_calorie, fill = poverty)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(mean_calorie, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            color = "black") +
  labs(
    x = "Region",
    y = "Calorie Per Capita",
    fill = "Status Kemiskinan"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "salmon"))
print(cal_chart)
```


```{r}
kp3_202303 <- kp3_202303 %>%
  mutate(poverty_ref = case_when(
    kapita < 0.8 * GK ~ 1,
    kapita >= 0.8 * GK & kapita < GK ~ 2,
    kapita >= GK & kapita < 1.2 * GK ~ 3,
    kapita >= 1.2 * GK & kapita < 1.6 * GK ~ 4,
    TRUE ~ 5
  ))

kp3_202303$poverty_ref <- factor(kp3_202303$poverty_ref, 
                          levels = c(1, 2, 3, 4, 5), 
                          labels = c("Sangat Miskin", "Miskin", "Hampir Miskin", "Rentan Miskin Lainnya", "Tidak Miskin"))


kp3_202303 <- kp3_202303 %>%
  mutate(island = case_when(
    r102 %in% c(9, 10, 11, 12, 13, 15, 18, 19) ~ "Flores",
    r102 %in% c(1, 2, 16, 17) ~ "Sumba",
    r102 %in% c(3, 4, 5, 6, 21, 71) ~ "Timor",
    r102 %in% c(7, 8, 14, 20) ~ "Lainnya",
    TRUE ~ "Unknown"
  )) %>%
  mutate(island = factor(island, levels = c("Flores", "Sumba", "Timor", "Lainnya")))

summary_data <- kp3_202303 %>%
  group_by(poverty_ref, kalori_kap, island) %>%
  summarise(size = sum(wert), .groups = 'drop') 

#bubble <- ggplot(summary_data, aes(x = size, y = kalori_kap, size = poverty_ref, color = island)) +
#  geom_point(alpha = 0.7) +
 # #scale_size_continuous(range = c(5, 20)) +  
  #labs(x = "Tingkatan Kemiskinan", y = "Konsumsi Kalori",
   #    size = "Ukuran", color = "Region",
    #   title = "Bubble Chart: Kemiskinan vs Konsumsi Kalori") +
#  theme_minimal() +
 # theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

#print(bubble)

```



#### d) Food Insecurity 

```{r,  echo=FALSE}
merged_kor2$food <- apply(merged_kor2[, c("R1704", "R1705", "R1706", "R1707", "R1708")], 1, function(x) ifelse(any(x == 1), 1, 0))

food_sec <- merged_kor2 %>%
  group_by(poverty, r105, food) %>%
  summarise(weighted_count = sum(FWT), .groups = 'drop') %>%
  group_by(poverty, r105) %>%
  mutate(total_weight = sum(weighted_count),
         percentage = (weighted_count / total_weight) * 100)

food_sec$r105 <- factor(food_sec$r105, levels = c(1, 2), labels = c("Pekotaan", "Perdesaan"))
food_sec$food <- factor(food_sec$food, levels = c(0, 1), labels = c("Tidak Rawan Pangan", "Rawan Pangan"))

food_secall <- merged_kor2 %>%
  group_by(poverty, food) %>%
  summarise(weighted_count = sum(FWT), .groups = 'drop') %>%
  group_by(poverty) %>%
  mutate(total_weight = sum(weighted_count),
         percentage = (weighted_count / total_weight) * 100)

food_secall$food <- factor(food_secall$food, levels = c(0, 1), labels = c("Tidak Rawan Pangan", "Rawan Pangan"))

combined_food <- bind_rows(food_sec, food_secall)
combined_food <- combined_food %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

f_chart <- ggplot(combined_food, aes(x = poverty, y = percentage, fill = food)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ r105) +
  labs(x = "", y = "Persentase", fill = "Status Rawan Pangan",
       title = "Persentase Status Rawan Pangan Berdasarkan Kemiskinan dan Region") +
  theme_minimal() +
  geom_text(aes(label = scales::percent(percentage / 100)), 
            position = position_dodge(width = 0.9), 
            size = 3,
            vjust = -0.5)

print(f_chart)

```


### 5. Asset Ownership

```{r,  echo=FALSE}
merged_kor2$asset <- apply(merged_kor2[, c("R2001A", "R2001B", "R2001C", "R2001D", "R2001E", "R2001F", "R2001G", "R2001H", "R2001I", "R2001J", "R2001K", "R2001L")], 1, function(x) ifelse(any(x == 1), 1, 0))

asset_own <- merged_kor2 %>%
  group_by(poverty, r105, asset) %>%
  summarise(weighted_count = sum(FWT), .groups = 'drop') %>%
  group_by(poverty, r105) %>%
  mutate(total_weight = sum(weighted_count),
         percentage = (weighted_count / total_weight) * 100)

asset_own$r105 <- factor(asset_own$r105, levels = c(1, 2), labels = c("Pekotaan", "Perdesaan"))
asset_own$food <- factor(asset_own$asset, levels = c(0, 1), labels = c("Tidak Memiliki Aset", "Memiliki Aset"))

asset_ownall <- merged_kor2 %>%
  group_by(poverty, asset) %>%
  summarise(weighted_count = sum(FWT), .groups = 'drop') %>%
  group_by(poverty) %>%
  mutate(total_weight = sum(weighted_count),
         percentage = (weighted_count / total_weight) * 100)

asset_ownall$food <- factor(asset_ownall$asset, levels = c(0, 1), labels = c("Tidak Memiliki Aset", "Memiliki Aset"))

combined_asset <- bind_rows(asset_own, asset_ownall)
combined_asset <- combined_asset %>%
  mutate(r105 = ifelse(is.na(r105), "Total", as.character(r105)))

a_chart <- ggplot(combined_asset, aes(x = poverty, y = percentage, fill = food)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ r105) +
  labs(x = "Poverty", y = "Persentase", fill = "Kepemilikan Aset",
       title = "Persentase Kepemilikan Berdasarkan Status Kemiskinan dan Daerah") +
  theme_minimal() +
  geom_text(aes(label = scales::percent(percentage / 100)), 
            position = position_dodge(width = 0.9), 
            size = 3,
            vjust = -0.5)

print(a_chart)
```


